---
title: "Dispos_2019_for_printouts"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(gsubfn)
library(lubridate)
library(openxlsx)
library(data.table)
library(lubridate)
library(readxl)
library(fuzzyjoin)
library(tidyverse)
library(tools)
library(vroom)
library(beepr)
library(sandwich) # for robust standard errors
library(lmtest) # to run tests using the robust standard errors
library(broom) # to get stuff out of the objects lmtest produces
library(pscl) #package for pseudo R2 (pR2 function)
library(RColorBrewer)
library(sjPlot)
library(waffle)
library(kableExtra)
library(webshot)
```

# Load and standardize collapsed dataset

Load collapsed dataset
```{r echo=FALSE, message=FALSE, warning=FALSE}
# load("G:/CrimJustice/Dispositions by Race and Ethnicity/DOJ Data 2019/Dispos.RData")
#load("G:/CrimJustice/Dispositions by Race and Ethnicity/DOJ Data 2019/Dispos Feb 2021.RData")
#load("G:/CrimJustice/Dispositions by Race and Ethnicity/DOJ Data 2019/Dispos Feb 17.RData") ## feb 17 version has incorrect hierarchy values, need to change max to min and re-run
load("G:/CrimJustice/Dispositions by Race and Ethnicity/DOJ Data 2019/Dispos Feb 24.RData")

df <- dispos %>%
  filter(arrest_felony_flag == 1) %>% #filter to only dispositions with a felony arrest, since there are also misdemeanor-only arrests for people who have other 2019 felony arrest dispositions
  #select(-cyc_order) %>%  #remove useless field
  rename(county = stp_ori_cnty_name,
         age = cyc_age)


rm(dispos)
```

Standardize factors
```{r echo=FALSE, message=FALSE, warning=FALSE}
#Standardize factors
df <- df %>%
 mutate(gender = as.factor(gender),
         
         # make age a number
         age = as.integer(age),
         age_first_dispo = as.integer(age_first_dispo),
        
         # add age category column
         age_cat = as.factor(case_when(
           age < 18 ~ "Under 18",
           age >= 18 & age <= 19 ~ "18-19 yrs",
           age > 19 & age <= 29 ~ "20-29 yrs",
           age > 29 & age <= 39 ~ "30-39 yrs",
           age > 39 & age <= 49 ~ "40-49 yrs",
           age > 49 & age <= 59 ~ "50-59 yrs",
           TRUE ~ "60+ yrs"
         )),
         
         
         # recode variables
         race = as.factor(case_when(
           race_code == "B" ~ "Black",
           race_code == "H" ~ "Hispanic",
           race_code == "W" ~ "White",
           race_code %in% c("A", "C", "D", "F", "G", "J", "K", "L", "P", "S", "U", "V", "Z") ~ "Asian/PI",
           race_code == "I" ~ "American Indian",
           T ~ "Other/Unknown")),
        
        # create a categorical variable for arrest offense type based only on felony arrest charges, prioritizing violent, then property, then drug
        arrest_f_offense_type = case_when(arrest_summ_f_violent_flag ==1 ~ "Violent",
                                         arrest_summ_f_property_flag == 1 ~ "Property",
                                         arrest_summ_f_drug_flag == 1 ~ "Drug",
                                         T ~ "Other"),
        
        #create combined suspended flag
        sent_suspended_flag = if_else(sent_ex_suspended_flag == 1 | sent_imp_suspended_flag ==1, 1, 0),
        
        #create a variable to track number of days since last dispo (will change to conviction when re-collapse completes)
        #think about what to do for ppl with no priors
        days_since_last_conv = as.integer(dispo_date - prior_conv_date), 
         
        #create scaled variables for hierarchy, and flip to make more intuitive so that higher values mean more severe crimes
        #for missing hierarchy values (represented by 1000000), set scaled value to 0
        prior_max_conv_hier_scaled = if_else(prior_max_conv_hier == 1000000, 0, -scale(prior_max_conv_hier)),
        max_arrest_hier_scaled = if_else(max_arrest_hier == 1000000, 0, -scale(max_arrest_hier)),
        max_court_hier_scaled = if_else(max_court_hier == 1000000, 0, -scale(max_court_hier)),
        max_conv_hier_scaled = if_else(max_conv_hier == 1000000, 0, -scale(max_conv_hier)),
        
        
        #create a dummy variable to indicate whether defendant has any priors, to use as an interaction term with days since last conviction
        prior_dummy = if_else(prior_convicted_flag_count >0, 1, 0),
        
        #create a categorical variable for prior record
        prior_record_cat = case_when(prior_sent_prison_flag_count >0  ~ "Prior prison",
                                      prior_sent_jail_flag_count >0  ~ "Prior jail with no prior prison",
                                      prior_convicted_flag_count >0 ~ "Prior conviction with no prior incarceration",
                                      T ~ "No prior convictions"),
        
        #create inverted variable with ceiling of years since prior conviction
        inv_yrs_since_prior_conv = if_else(prior_convicted_flag_count >0, 1/ceiling(days_since_last_conv/365.25), 0),
        
        #create years of prior prison and jail for ease of interpretability
        years_prior_prison = days_prior_prison/365.25,
        years_prior_jail = days_prior_jail/365.25
        
        )


#set base levels for regressions
df$race <- relevel(df$race, ref="White")
df$gender <- relevel(df$gender, ref="M")
df$arrest_f_offense_type <- as.factor(df$arrest_f_offense_type)
df$arrest_f_offense_type <- relevel(df$arrest_f_offense_type, ref="Property")
df$county <- as.factor(df$county)
df$county <- relevel(df$county, ref ="LOS ANGELES")


```

# Load and tidy CA population data from DOF
tidy ca_age data
```{r}
ca_age <- read_xlsx("G:/CrimJustice/Dispositions by Race and Ethnicity/DOJ Data 2019/DOF demographic data 2019/P1B_State_Age.xlsx", sheet = 3, skip = 2, col_names =T) 


# ca_age_full <- ca_age %>%
#   # extract columns
#    select(`Population`, `2019`) %>%
#   
#   # rename column names
#   rename(Age = Population) %>%
#   
#   # remove rows that are not ages
#   slice(-c(1, 103)) %>%
# 
#   # recode variable
#   mutate(`2019` = as.integer(`2019`),
#           Age = as.integer(if_else(Age == "100+", "100", Age)),
#          
#          # add age category column
#          `Age Category` = as.factor(case_when(
#            Age <= 14 ~ "0-14 yrs",
#            Age > 14 & Age <= 19 ~ "15-19 yrs",
#            Age > 19 & Age <= 29 ~ "20-29 yrs",
#            Age > 29 & Age <= 39 ~ "30-39 yrs",
#            Age > 39 & Age <= 49 ~ "40-49 yrs",
#            Age > 49 & Age <= 59 ~ "50-59 yrs",
#            TRUE ~ "60+ yrs"
#          ))) 

ca_age18 <- ca_age %>%
 # extract columns
   select(`Population`, `2019`) %>%
  
  # rename column names
  rename(Age = Population) %>%
  
  # remove rows that are not ages
  slice(-c(1, 103)) %>%

  # recode variable
  mutate(`2019` = as.integer(`2019`),
          Age = as.integer(if_else(Age == "100+", "100", Age)),
         
         # add age category column
         `Age Category` = as.factor(case_when(
           Age <= 17 ~ "0-17 yrs",
           Age > 17 & Age <= 19 ~ "18-19 yrs",
           Age > 19 & Age <= 29 ~ "20-29 yrs",
           Age > 29 & Age <= 39 ~ "30-39 yrs",
           Age > 39 & Age <= 49 ~ "40-49 yrs",
           Age > 49 & Age <= 59 ~ "50-59 yrs",
           TRUE ~ "60+ yrs"
         ))) %>%
  
  #remove people less than 18
  filter(`Age Category`!= "0-17 yrs") %>%
  
  # add frequency column
  group_by(`Age Category`) %>%
  summarise(Frequency = sum(`2019`)) %>%
  
  # add percentage column
  mutate(Percentage = Frequency / sum(Frequency) * 100)

# ca_age <- ca_age %>%
#   # extract columns
#    select(`Population`, `2019`) %>%
#   
#   # rename column names
#   rename(Age = Population) %>%
#   
#   # remove rows that are not ages
#   slice(-c(1, 103)) %>%
# 
#   # recode variable
#   mutate(`2019` = as.integer(`2019`),
#           Age = as.integer(if_else(Age == "100+", "100", Age)),
#          
#          # add age category column
#          `Age Category` = as.factor(case_when(
#            Age <= 14 ~ "0-14 yrs",
#            Age > 14 & Age <= 19 ~ "15-19 yrs",
#            Age > 19 & Age <= 29 ~ "20-29 yrs",
#            Age > 29 & Age <= 39 ~ "30-39 yrs",
#            Age > 39 & Age <= 49 ~ "40-49 yrs",
#            Age > 49 & Age <= 59 ~ "50-59 yrs",
#            TRUE ~ "60+ yrs"
#          ))) %>%
#   
#   # add frequency column
#   group_by(`Age Category`) %>%
#   summarise(Frequency = sum(`2019`)) %>%
#   
#   # add percentage column
#   mutate(Percentage = Frequency / sum(Frequency) * 100)
ca_age18
```

tidy ca_eth data
```{r}
ca_eth <- read_xlsx("G:/CrimJustice/Dispositions by Race and Ethnicity/DOJ Data 2019/DOF demographic data 2019/P1D_State_Race_Ethnicity.xlsx", sheet = 2, skip = 2, col_names =T) 

ca_eth <- ca_eth %>%
  # extract columns
   select(`Race/Ethnicity`, `2019`) %>%
  
  # remove last two rows which are total and a note
  slice(-c(8,9)) %>%

  # recode variable
  mutate(`2019` = as.integer(`2019`)) %>%
  # add percentage column
  mutate(Percentage = `2019` / sum(`2019`) * 100)

ca_eth
  
```

tidy ca_sex data
```{r}
ca_sex <- read_xlsx("G:/CrimJustice/Dispositions by Race and Ethnicity/DOJ Data 2019/DOF demographic data 2019/P1C_State_Age-Group_Sex.xlsx", sheet = 2, skip = 2, col_names =T) 

ca_sex <- ca_sex %>%
  # extract columns
   select(`Women, by age group:`, `2019`) %>%
  
   # rename column names
  rename(Sex = `Women, by age group:`) %>%
  
  #select only adult ages
  slice(5:22, 30:47 ) %>%
  
  rename(age = Sex) %>%
  rename(n = `2019`) %>%
  
  mutate(sex = c(rep("Women", 18), rep("Men", 18)),
         n = as.numeric(n)) %>%
  
  #pivot_wider(names_from = sex, values_from = n) %>%
  
  group_by(sex) %>%
  summarise(total = sum(n)) %>%
  mutate(percent = total/sum(total)*100)
  
  
  # select rows that are total women and total men
  #slice(23,48) %>%
  
# 
#   # recode variables
#   mutate(`2019` = as.integer(`2019`),
#          Sex = if_else(Sex == "Total women", "Women", "Men")) %>%
#   
#   # add percentage column
#   mutate(Percentage = `2019` / sum(`2019`) * 100)
  

ca_sex
```



# Descriptives

Numbers for Figure 1 flow chart
```{r warning=FALSE}
#all
nrow(df)

df %>%
  count(court_disp) %>%
  mutate(percent = n/sum(n)) #final court dispositions vs law enforcement/prosecution release dispositions

df %>%
  filter(court_disp ==1) %>%
  count(convicted_flag) %>%
  mutate(percent = n/sum(n)) #convicted out of those with a court disposition

# df %>%
#   filter(convicted_flag ==1) %>%
#   count(sent_prison_flag) %>%
#   mutate(percent = n/sum(n)) #prison sentences out of those convicted

sent = df %>%
  filter(convicted_flag ==1) %>%
  count(sent_prison_flag, sent_jail_flag, sent_probation_flag) %>%
  mutate(percent = n/sum(n)) #jail and probation sentences out of those convicted

write.csv(sent, "sentencing_table.csv")
```

Demographics of felony defendants (court disposition)
```{r message=FALSE, warning=FALSE}
df %>%
  filter(court_disp ==1) %>%
  count(age_cat) %>%
  filter(age_cat != "Under 18") %>%
  mutate(percent = round(n/sum(n)*100, 1)) #age distribution by age category

df %>%
  filter(court_disp ==1) %>%
  count(gender) %>%
  filter(gender !="X") %>%
  mutate(percent = round(n/sum(n)*100, 1)) #gender distribution

#race distribution filings
df %>%
  filter(court_disp ==1) %>%
  count(race) %>%
  mutate(percent = round(n/sum(n)*100, 1)) #race distribution

#race distribution all
df %>%
  count(race) %>%
  mutate(percent = round(n/sum(n)*100, 1)) #race distribution

aggregate(age ~ race, data = df %>% filter(court_disp ==1), FUN = mean)

df %>%
  filter(court_disp ==1) %>%
  group_by(race, age_cat) %>%
  summarise(n = n()) %>%
  group_by(race) %>%
  mutate(percent = round(n/sum(n)*100, 1)) %>%
  ungroup() %>%
  select(-n) %>%
  pivot_wider(names_from = age_cat, values_from = percent) %>%
  select(race, 'Under 18', '18-19 yrs', '20-29 yrs', '30-39 yrs', '40-49 yrs', '50-59 yrs', '60+ yrs')


#the citizen flag must be horribly inaccurate, only capturing a very small percentage of individuals
df %>%
  filter(court_disp ==1) %>%
  group_by(race, citizen) %>%
  summarise(n = n()) %>%
  group_by(race) %>%
  mutate(percent = round(n/sum(n)*100, 1)) %>%
  filter(citizen ==1)

```

Prior record of felony defendants (court disposition)
```{r message=FALSE, warning=FALSE}
df %>%
  filter(court_disp ==1) %>%
  count(prior_sent_prison_flag_count >0) %>%
  mutate(percent = round(n/sum(n)*100, 1)) # prior prison

df %>%
  filter(court_disp ==1) %>%
  count(prior_sent_prison_flag_count == 0 & prior_sent_jail_flag_count >0) %>%
  mutate(percent = round(n/sum(n)*100, 1)) #prior jail with no prior prison

df %>%
  filter(court_disp ==1) %>%
  count(prior_sent_prison_flag_count == 0 & prior_sent_jail_flag_count == 0 & prior_sent_probation_flag_count >0) %>%
  mutate(percent = round(n/sum(n)*100, 1)) #prior probation with no prior incarceration

df %>%
  filter(court_disp ==1) %>%
  count(prior_sent_prison_flag_count == 0 & prior_sent_jail_flag_count == 0 & prior_convicted_flag_count >0) %>%
  mutate(percent = round(n/sum(n)*100, 1)) #prior conviction with no prior incarceration

df %>%
  filter(court_disp ==1) %>%
  count(prior_convicted_flag_count ==0) %>%
  mutate(percent = round(n/sum(n)*100, 1)) #no prior convictions

df %>%
  filter(court_disp ==1) %>%
  count(prior_conviction_felony_flag_count >0) %>%
  mutate(percent = round(n/sum(n)*100, 1)) # has prior felony convictions

df %>%
  filter(court_disp ==1) %>%
  count(prior_conviction_felony_flag_count == 0 & prior_conviction_misd_flag_count >0) %>%
  mutate(percent = round(n/sum(n)*100, 1)) # has prior misdemeanors with no prior felonies

df %>%
  filter(court_disp ==1) %>%
  count(prior_conviction_felony_flag_count) %>%
  mutate(percent = round(n/sum(n)*100, 1)) # number of prior felony convictions

df %>%
  filter(court_disp ==1) %>%
  count(prior_conviction_misd_flag_count) %>%
  mutate(percent = round(n/sum(n)*100, 1)) # number of prior misdo convictions

mean(df$prior_conviction_felony_flag_count) #mean prior felonies
mean(df$prior_conviction_misd_flag_count) #mean prior misdos



#### prior sentences

df %>%
  filter(court_disp == 1) %>%
  mutate(prior_record_cat = case_when(prior_sent_prison_flag_count >0  ~ "Prior prison",
                                      prior_sent_jail_flag_count >0 & days_prior_jail >=365 ~ "Prior jail, 1 year or more",
                                      prior_sent_jail_flag_count >0  ~ "Prior jail, less than 1 year",
                                      prior_convicted_flag_count >0 ~ "Prior conviction with no prior incarceration",
                                      T ~ "No prior convictions")) %>%
  count(prior_record_cat) %>%
  mutate(percent = round(n/sum(n)*100, 1))


# mean days prior prison, for those who have had prior prison
df %>%
  filter(prior_sent_prison_flag_count >0) %>%
  summarise(mean(days_prior_prison, na.rm = T)) #mean prior prison time is nearly 8 years, ouch.
  # count(is.na(days_prior_prison)) %>%
  # mutate(percent = round(n/sum(n)*100, 1))

#mean days prior jail, for those who have had prior jail
df %>%
  filter(prior_sent_jail_flag_count >0) %>%
  summarise(mean(days_prior_jail, na.rm = T)) # mean prior jail time is a little less than 2 years.
  # count(is.na(days_prior_jail)) %>%
  # mutate(percent = round(n/sum(n)*100, 1))

### categorical priors
df %>%
  filter(court_disp ==1) %>%
  count(prior_record_cat) %>%
  mutate(percent = round(n/sum(n)*100, 1))
```

Arrest offense type for felony defendants (court dispo)
```{r message=FALSE, warning=FALSE}

df %>%
  filter(court_disp ==1) %>%
  count(arrest_f_offense_type) %>%
  mutate(percent = round(n/sum(n)*100, 1)) #arrest offense type for felony arrest offenses only



```

Descriptive info on sentences
```{r message=FALSE, warning=FALSE}
df %>%
  filter(sent_prison_flag ==1) %>%
  summarise(mean(prison_length_days, na.rm = T)) #mean prison sentence term is 2039 days (5.6 years)

df %>%
  filter(sent_jail_flag == 1) %>%
  summarise(mean(jail_length_days, na.rm = T)) #mean jail sentence term is 272 days (0.75 years)

df %>%
  filter(sent_probation_flag == 1) %>%
  #count(probation_length_days)
  summarise(mean(probation_length_days, na.rm = T)) #mean probation sentence term is 1242 days (3.4 years). this might be thrown off by suspended sentences that lack probation length (many of these are zeros), and some wonky extremely large term lengths (there is one over 100,000 days = 273 years.... seems excessive)


```

Descriptive info on priors by race and age category
```{r message=FALSE, warning=FALSE}

#priors by race
df %>%
  group_by(race) %>%
  summarise(n = n(), mean_age = mean(age), mean_prior_fs = mean(prior_conviction_felony_flag_count), mean_prior_ms = mean(prior_conviction_misd_flag_count))

df %>%
  group_by(age_cat) %>%
  summarise(n = n(), mean_prior_fs = mean(prior_conviction_felony_flag_count), mean_prior_ms = mean(prior_conviction_misd_flag_count))

df %>%
  filter(court_disp == 1) %>%
  mutate(prior_f_violent = if_else(prior_conviction_summ_f_violent_flag_count > 0, 1, 0),
         prior_m_violent = if_else(prior_conviction_summ_m_violent_flag_count > 0, 1, 0),
         prior_f_drug = if_else(prior_conviction_summ_f_drug_flag_count > 0, 1, 0),
         prior_m_drug = if_else(prior_conviction_summ_m_drug_flag_count > 0, 1, 0),
         prior_f_prop = if_else(prior_conviction_summ_f_property_flag_count > 0, 1, 0),
         prior_m_prop = if_else(prior_conviction_summ_m_property_flag_count > 0, 1, 0)) %>%
  group_by(race) %>%
  summarise(across(c(prior_f_violent, prior_m_violent, prior_f_drug, prior_m_drug, prior_f_prop, prior_m_prop), mean))
  #summarise(mean(prior_f_violent), mean(prior_m_violent), mean(prior_f_drug), mean(prior_m_drug), mean(prior_f_prop), mean(prior_m_prop))

  
df %>%
  filter(court_disp == 1) %>%
  group_by(race) %>%
  summarise(across(c(prior_conviction_summ_f_violent_flag_count, prior_conviction_summ_m_violent_flag_count, prior_conviction_summ_f_drug_flag_count, prior_conviction_summ_m_drug_flag_count, prior_conviction_summ_f_property_flag_count, prior_conviction_summ_m_property_flag_count, prior_conviction_summ_f_other_sex_flag_count, prior_conviction_summ_m_other_sex_flag_count, prior_conviction_summ_f_other_flag_count, prior_conviction_summ_m_other_flag_count), ~mean(.x>0)))
#biggest racial differences in: prior felony violent, prior misdo drug


#prior convictions, at least one by type, by race
df %>%
  filter(prior_arrest_summ_f_drug_flag_count >0) %>%
  group_by(race) %>%
  summarise(across(matches("prior_conviction_summ"), ~mean(.x>0)))


#prior drug by race
df %>%
  filter(court_disp == 1) %>%
  mutate(prior_drug = prior_conviction_summ_f_drug_flag_count + prior_conviction_summ_m_drug_flag_count) %>%
  group_by(race) %>%
  summarise(mean(prior_drug>0))
  
#time since prior conviction by race
df %>%
  filter(court_disp == 1) %>%
  group_by(race) %>%
  summarise(mean(days_since_last_conv))

#age at first arrest by race
df %>%
  filter(court_disp == 1) %>%
  group_by(race) %>%
  summarise(quantile(age_first_dispo, probs = c(0.25)), mean(age_first_dispo), median(age_first_dispo), quantile(age_first_dispo, probs = c(0.75)))
```

Arrest offense categories and counts by race
```{r message=FALSE, warning=FALSE}
#arrest offense type by race
df %>%
  filter(court_disp == 1) %>%
  select(-contains("prior")) %>%
  group_by(race) %>%
  summarise(across(contains("arrest_summ"),  ~round(mean(.x>0)*100,1)))

#using categorical felony arrest offense type variable
df %>%
  filter(court_disp == 1) %>%
  group_by(race) %>%
  count(arrest_f_offense_type) %>%
  mutate(percent = round(n/sum(n)*100, 1)) %>%
  select(-n) %>%
  pivot_wider(names_from = arrest_f_offense_type, values_from = percent) %>%
  select(race, Violent, Drug, Property, Other)


#offense categories by race
df %>%
  group_by(race) %>%
  summarise(violentf = mean(arrest_f_offense_type == "Violent"), propf = mean(arrest_f_offense_type == "Property"), drugf = mean(arrest_f_offense_type == "Drug"), otherf = mean(arrest_f_offense_type == "Other"))

#offense categories by age category
df %>%
  group_by(age_cat) %>%
  summarise(violentf = mean(arrest_f_offense_type == "Violent"), propf = mean(arrest_f_offense_type == "Property"), drugf = mean(arrest_f_offense_type == "Drug"), otherf = mean(arrest_f_offense_type == "Other"))

#prior offense category counts by race
df %>%
  group_by(race) %>%
  summarise(prior_violentf = mean(prior_arrest_summ_f_violent_flag_count), prior_propf = mean(prior_arrest_summ_f_property_flag_count), prior_drugf = mean(prior_arrest_summ_f_drug_flag_count), prior_otherf = mean(prior_arrest_summ_f_other_flag_count))


#number of arrest charges and offense hierarchy by race
df %>%
  group_by(race) %>%
  summarise(fcharge = mean(arrested_fcharge_count), mcharge = mean(arrested_mcharge_count), arresthier = mean(max_arrest_hier))
```

Descriptives of outcomes for regressions
```{r message=FALSE, warning=FALSE}
#filing rate and overall conviction rate out of all arrests by race
df %>%
  group_by(race) %>%
  summarise(filing = round(mean(filed_flag)*100,1), convictions_all = round(mean(convicted_flag)*100, 1))

#conviction rate out of cases filed by race
df %>%
  filter(filed_flag == 1) %>%
  group_by(race) %>%
  summarise(conviction_of_filings = round(mean(convicted_flag)*100,1))

#felony conviction rate out of cases filed, all
df %>%
  filter(convicted_flag == 1) %>%
  count(conviction_felony_flag, conviction_misd_flag, conviction_inf_flag) %>%
  mutate(perc = round(n/sum(n)*100, 1))


#felony conviction rate out of cases filed by race
df %>%
  filter(convicted_flag == 1) %>%
  group_by(race) %>%
  summarise(felony_conv_of_conv = round(mean(conviction_felony_flag)*100,1))


#prison sentencing rate out of convicted cases by race
df %>%
  filter(convicted_flag == 1) %>%
  group_by(race) %>%
  summarise(prison_of_convicted = round(mean(sent_prison_flag)*100,1))

#prison sentencing rate out of felony convicted cases by race
df %>%
  filter(convicted_flag == 1, conviction_felony_flag ==1) %>%
  group_by(race) %>%
  summarise(prison_of_fel_convicted = round(mean(sent_prison_flag)*100,1))


#prison sentence length out of cases sentenced to prison
df %>%
  filter(sent_prison_flag == 1) %>%
  group_by(race) %>%
  summarise(prison_length_yrs = round(mean(prison_length_days, na.rm = T)/365.25, 1), perc_length_na = round(mean(is.na(prison_length_days))*100,1))

```

Charts for Figure 6 -- outcomes by race
```{r}
#conviction rate out of cases filed by race
df %>%
  filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(race) %>%
  summarise(conviction_of_filings = round(mean(convicted_flag)*100,1)) %>%
  ggplot(aes(fill=race, y= conviction_of_filings, x=race)) + 
    geom_bar(position="dodge", stat="identity") +
    scale_color_brewer(palette = "Set1")+
    ylim(0, 100) +
    geom_text(aes(label=round(conviction_of_filings, 0)), position=position_dodge(width=0.9), vjust=-0.25)

ggsave("conviction rate by race.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")

#felony conviction rate out of cases filed by race
df %>%
  filter(convicted_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(race) %>%
  summarise(felony_conv_of_conv = round(mean(conviction_felony_flag)*100,1)) %>%
  ggplot(aes(fill=race, y= felony_conv_of_conv, x=race)) + 
    geom_bar(position="dodge", stat="identity") +
    scale_color_brewer(palette = "Set1")+
    ylim(0, 100) +
    geom_text(aes(label=round(felony_conv_of_conv, 0)), position=position_dodge(width=0.9), vjust=-0.25)

ggsave("felony conviction rate by race.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")


#prison sentencing rate out of felony convicted cases by race
df %>%
  filter(conviction_felony_flag ==1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(race) %>%
  summarise(prison_of_fel_convicted = round(mean(sent_prison_flag)*100,1)) %>%
  ggplot(aes(fill=race, y= prison_of_fel_convicted, x=race)) + 
    geom_bar(position="dodge", stat="identity") +
    scale_color_brewer(palette = "Set1")+
    ylim(0, 100) +
    geom_text(aes(label=round(prison_of_fel_convicted, 0)), position=position_dodge(width=0.9), vjust=-0.25)

ggsave("prison sentencing rate by race.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")


#prison sentence length out of cases sentenced to prison
df %>%
  filter(sent_prison_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(race) %>%
  summarise(prison_length_yrs = round(mean(prison_length_days, na.rm = T)/365.25, 1), perc_length_na = round(mean(is.na(prison_length_yrs))*100,1)) %>%
  ggplot(aes(fill=race, y= prison_length_yrs, x=race)) + 
    geom_bar(position="dodge", stat="identity") +
    scale_color_brewer(palette = "Set1")+
    ylim(0, 15) +
    geom_text(aes(label=round(prison_length_yrs, 1)), position=position_dodge(width=0.9), vjust=-0.25)

ggsave("prison sentence length by race.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")


```

Charts for Figure 6 -- outcomes by criminal history
```{r}
#conviction rate out of cases filed by prior_record_cat
df %>%
  filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(prior_record_cat) %>%
  summarise(conviction_of_filings = round(mean(convicted_flag)*100,1)) %>%
  ggplot(aes(fill=prior_record_cat, y= conviction_of_filings, x=prior_record_cat)) + 
    geom_bar(position="dodge", stat="identity") +
    scale_color_brewer(palette = "Set1")+
    theme(axis.text.x=element_blank()) +
    ylim(0, 100) +
    geom_text(aes(label=round(conviction_of_filings, 0)), position=position_dodge(width=0.9), vjust=-0.25)

ggsave("conviction rate by prior record.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")

#felony conviction rate out of cases filed by prior_record_cat
df %>%
  filter(convicted_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(prior_record_cat) %>%
  summarise(felony_conv_of_conv = round(mean(conviction_felony_flag)*100,1)) %>%
  ggplot(aes(fill=prior_record_cat, y= felony_conv_of_conv, x=prior_record_cat)) + 
    geom_bar(position="dodge", stat="identity") +
    scale_color_brewer(palette = "Set1")+
    theme(axis.text.x=element_blank()) +
    ylim(0, 100) +
    geom_text(aes(label=round(felony_conv_of_conv, 0)), position=position_dodge(width=0.9), vjust=-0.25)

ggsave("felony conviction rate by prior record.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")

#prison sentencing rate out of felony convicted cases by prior_record_cat
df %>%
  filter(conviction_felony_flag ==1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(prior_record_cat) %>%
  summarise(prison_of_fel_convicted = round(mean(sent_prison_flag)*100,1)) %>%
  ggplot(aes(fill=prior_record_cat, y= prison_of_fel_convicted, x=prior_record_cat)) + 
    geom_bar(position="dodge", stat="identity") +
    scale_color_brewer(palette = "Set1")+
    theme(axis.text.x=element_blank()) +
    ylim(0, 100) +
    geom_text(aes(label=round(prison_of_fel_convicted, 0)), position=position_dodge(width=0.9), vjust=-0.25)

ggsave("prison sentencing rate by prior record.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")

#prison sentence length out of cases sentenced to prison
df %>%
  filter(sent_prison_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(prior_record_cat) %>%
  summarise(prison_length_yrs = round(mean(prison_length_days, na.rm = T)/365.25, 1), perc_length_na = round(mean(is.na(prison_length_yrs))*100,1)) %>%
  ggplot(aes(fill=prior_record_cat, y= prison_length_yrs, x=prior_record_cat)) + 
    geom_bar(position="dodge", stat="identity") +
    scale_color_brewer(palette = "Set1")+
    theme(axis.text.x=element_blank()) +
    ylim(0, 15) +
    geom_text(aes(label=round(prison_length_yrs, 1)), position=position_dodge(width=0.9), vjust=-0.25)

ggsave("prison sentence length by prior record.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")


```

Charts for Figure 6 -- outcomes by felony arrest offense type
```{r}
#conviction rate out of cases filed by arrest_f_offense_type
df %>%
  filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(arrest_f_offense_type) %>%
  summarise(conviction_of_filings = round(mean(convicted_flag)*100,1)) %>%
  ggplot(aes(fill=arrest_f_offense_type, y= conviction_of_filings, x=arrest_f_offense_type)) + 
    geom_bar(position="dodge", stat="identity") +
    scale_color_brewer(palette = "Set1")+
    ylim(0, 100) +
    geom_text(aes(label=round(conviction_of_filings, 0)), position=position_dodge(width=0.9), vjust=-0.25)

ggsave("conviction rate by arrest offense type.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")


#felony conviction rate out of cases filed by arrest_f_offense_type
df %>%
  filter(convicted_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(arrest_f_offense_type) %>%
  summarise(felony_conv_of_conv = round(mean(conviction_felony_flag)*100,1)) %>%
  ggplot(aes(fill=arrest_f_offense_type, y= felony_conv_of_conv, x=arrest_f_offense_type)) + 
    geom_bar(position="dodge", stat="identity") +
    scale_color_brewer(palette = "Set1")+
    ylim(0, 100) +
    geom_text(aes(label=round(felony_conv_of_conv, 0)), position=position_dodge(width=0.9), vjust=-0.25)

ggsave("felony conviction rate by arrest offense type.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")


#prison sentencing rate out of felony convicted cases by arrest_f_offense_type
df %>%
  filter(conviction_felony_flag ==1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(arrest_f_offense_type) %>%
  summarise(prison_of_fel_convicted = round(mean(sent_prison_flag)*100,1)) %>%
  ggplot(aes(fill=arrest_f_offense_type, y= prison_of_fel_convicted, x=arrest_f_offense_type)) + 
    geom_bar(position="dodge", stat="identity") +
    scale_color_brewer(palette = "Set1")+
    ylim(0, 100) +
    geom_text(aes(label=round(prison_of_fel_convicted, 0)), position=position_dodge(width=0.9), vjust=-0.25)

ggsave("prison sentencing rate by arrest offense type.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")


#prison sentence length out of cases sentenced to prison
df %>%
  filter(sent_prison_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(arrest_f_offense_type) %>%
  summarise(prison_length_yrs = round(mean(prison_length_days, na.rm = T)/365.25, 1), perc_length_na = round(mean(is.na(prison_length_yrs))*100,1)) %>%
  ggplot(aes(fill=arrest_f_offense_type, y= prison_length_yrs, x=arrest_f_offense_type)) + 
    geom_bar(position="dodge", stat="identity") +
    scale_color_brewer(palette = "Set1")+
    ylim(0, 15) +
    geom_text(aes(label=round(prison_length_yrs, 1)), position=position_dodge(width=0.9), vjust=-0.25)

ggsave("prison sentence length by arrest offense type.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")


```


Charts for Appendix D -- Conviction rates
```{r echo=FALSE, fig.height=10, fig.width=8, message=FALSE, warning=FALSE}


#Conviction rate by arrest offense type, prior record, and race
df %>%
  filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F"),
         age >= 18) %>%
  filter(arrest_f_offense_type %in% c("Violent", "Property", "Drug")) %>%
  group_by(race, prior_record_cat, arrest_f_offense_type) %>%
  summarise(mean(convicted_flag)) %>%
  rename(conviction_rate = 'mean(convicted_flag)') %>%
  mutate(conviction_rate = conviction_rate*100) %>%
  ggplot(aes(fill=race, y= conviction_rate, x= prior_record_cat)) + 
    facet_grid(rows = vars(arrest_f_offense_type)) +
    geom_bar(position=position_dodge(width=0.9), stat="identity", width = 0.85) +
    scale_color_brewer(palette = "Set1") +
    #theme(axis.text.x = element_text(angle = 90))+
    geom_text(aes(label=round(conviction_rate, 0)), position=position_dodge(width=0.9), vjust=-0.25) +
    ylim(0, 100) +
    scale_x_discrete(labels = c("No prior\n convictions", "Prior conviction\n with no prior\n incarceration", "Prior jail\n with no prior\n prison", "Prior prison"))+
    xlab("Prior criminal record") +
    ylab("Conviction rate")

ggsave("appendix D - conviction rate by arrest offense type, prior record, and race.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")

```

Charts for Appendix D -- Felony conviction rates
```{r echo=FALSE, fig.height=10, fig.width=8,message=FALSE, warning=FALSE}

#Felony conviction rate by arrest offense type and race
df %>%
  filter(convicted_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F"),
         age >= 18) %>%
  filter(arrest_f_offense_type %in% c("Violent", "Property", "Drug")) %>%
  group_by(race, prior_record_cat, arrest_f_offense_type) %>%
  summarise(mean(conviction_felony_flag)) %>%
  rename(felony_conviction_rate = 'mean(conviction_felony_flag)') %>%
  mutate(felony_conviction_rate = felony_conviction_rate*100) %>%
  ggplot(aes(fill=race, y= felony_conviction_rate, x= prior_record_cat)) + 
    facet_grid(rows = vars(arrest_f_offense_type)) +
    geom_bar(position=position_dodge(width=0.9), stat="identity", width = 0.85) +
    scale_color_brewer(palette = "Set1") +
    #theme(axis.text.x = element_text(angle = 90))+
    geom_text(aes(label=round(felony_conviction_rate, 0)), position=position_dodge(width=0.9), vjust=-0.25) +
    ylim(0, 100) +
    scale_x_discrete(labels = c("No prior\n convictions", "Prior conviction\n with no prior\n incarceration", "Prior jail\n with no prior\n prison", "Prior prison"))+
    xlab("Prior criminal record") +
    ylab("Felony conviction rate")

ggsave("appendix D - felony conviction rate by arrest offense type, prior record, and race.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")

```

Charts for Appendix D -- Prison sentencing rates
```{r echo=FALSE, fig.height=10, fig.width=8,message=FALSE, warning=FALSE}

#Prison sentencing rate of the felony convicted by arrest offense type and race
df %>%
  filter(conviction_felony_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F"),
         age >= 18) %>%
  filter(arrest_f_offense_type %in% c("Violent", "Property", "Drug")) %>%
  group_by(race, prior_record_cat, arrest_f_offense_type) %>%
  summarise(mean(sent_prison_flag)) %>%
  rename(prison_sent_rate = 'mean(sent_prison_flag)') %>%
  mutate(prison_sent_rate = prison_sent_rate*100) %>%
  ggplot(aes(fill=race, y= prison_sent_rate, x= prior_record_cat)) + 
    facet_grid(rows = vars(arrest_f_offense_type)) +
    geom_bar(position=position_dodge(width=0.9), stat="identity", width = 0.85) +
    scale_color_brewer(palette = "Set1") +
    #theme(axis.text.x = element_text(angle = 90))+
    geom_text(aes(label=round(prison_sent_rate, 0)), position=position_dodge(width=0.9), vjust=-0.25) +
    ylim(0, 100) +
    scale_x_discrete(labels = c("No prior\n convictions", "Prior conviction\n with no prior\n incarceration", "Prior jail\n with no prior\n prison", "Prior prison"))+
    xlab("Prior criminal record") +
    ylab("Prison sentencing rate")

ggsave("appendix D - prison rate by arrest offense type, prior record, and race.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")


```

Charts for Appendix D -- Prison sentence length
```{r echo=FALSE, fig.height=10, fig.width=8,message=FALSE, warning=FALSE}

#Prison sentencing length of the prison sentenced by arrest offense type and race
df %>%
  filter(sent_prison_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F"),
         age >= 18) %>%
  filter(arrest_f_offense_type %in% c("Violent", "Property", "Drug")) %>%
  group_by(race, prior_record_cat, arrest_f_offense_type) %>%
  summarise(mean(prison_length_days/365.25, na.rm = T)) %>%
  rename(prison_length = 'mean(prison_length_days/365.25, na.rm = T)') %>%
  ggplot(aes(fill=race, y= prison_length, x= prior_record_cat)) + 
    facet_grid(rows = vars(arrest_f_offense_type)) +
    geom_bar(position=position_dodge(width=0.9), stat="identity", width = 0.85) +
    scale_color_brewer(palette = "Set1") +
    #theme(axis.text.x = element_text(angle = 90))+
    geom_text(aes(label=round(prison_length, 1)), position=position_dodge(width=0.9), vjust=-0.25) +
    ylim(0, 16) +
    scale_x_discrete(labels = c("No prior\n convictions", "Prior conviction\n with no prior\n incarceration", "Prior jail\n with no prior\n prison", "Prior prison")) +
    labs(x = "Prior criminal record", y = "Prison sentence length (years)", fill = "Race")


ggsave("appendix D - prison length by arrest offense type, prior record, and race.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")



```


Numbers for Appendix C
```{r}
#for C1
df %>%
  filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F"),
         age >= 18) %>%
  group_by(race) %>%
  summarise(conviction_rate = round(mean(convicted_flag)*100, 1))


#for C2
df %>%
  filter(convicted_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F"),
         age >= 18) %>%
  group_by(race) %>%
  summarise(prison_rate = round(mean(sent_prison_flag)*100, 1))
```

Appendix Table B1
```{r}
#conviction rate out of cases filed by race
df %>%
  filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(race) %>%
  summarise(conviction_of_filings = round(mean(convicted_flag)*100,1),
            non_conviction_of_filings = 100-conviction_of_filings) %>%
  pivot_longer(cols = c(conviction_of_filings, non_conviction_of_filings), names_to = "var") %>%
  pivot_wider(names_from = race, values_from = value) %>%
  bind_cols(total = df %>%
    filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
      summarise(conviction_of_filings = round(mean(convicted_flag)*100,1),
            non_conviction_of_filings = 100-conviction_of_filings) %>%
      pivot_longer(cols = c(conviction_of_filings, non_conviction_of_filings), names_to = "variable") %>%
      rename(Total = value)) %>%
  select(var, Total, `Asian/PI`, "Black", "Hispanic", "White") %>%

bind_rows(
#felony conviction rate out of cases filed by race
df %>%
  filter(convicted_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(race) %>%
  summarise(felony_conv_of_conv = round(mean(conviction_felony_flag)*100,1),
            non_felony_conv_of_conv = 100-felony_conv_of_conv) %>%
  pivot_longer(cols = c(felony_conv_of_conv, non_felony_conv_of_conv), names_to = "var") %>%
  pivot_wider(names_from = race, values_from = value) %>%
  bind_cols(total = df %>%
    filter(convicted_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
      summarise(felony_conv_of_conv = round(mean(conviction_felony_flag)*100,1),
            non_felony_conv_of_conv = 100-felony_conv_of_conv) %>%
      pivot_longer(cols = c(felony_conv_of_conv, non_felony_conv_of_conv), names_to = "variable") %>%
      rename(Total = value))%>%
  select(var, Total, `Asian/PI`, "Black", "Hispanic", "White")
) %>%
  
bind_rows(

#prison sentencing rate out of felony convicted cases by race
df %>%
  filter(conviction_felony_flag ==1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(race) %>%
  summarise(prison_of_fel_convicted = round(mean(sent_prison_flag)*100,1),
            non_prison_of_fel_convicted = 100-prison_of_fel_convicted) %>%
  pivot_longer(cols = c(prison_of_fel_convicted, non_prison_of_fel_convicted), names_to = "var") %>%
  pivot_wider(names_from = race, values_from = value) %>%
  bind_cols(total = df %>%
    filter(conviction_felony_flag ==1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
      summarise(prison_of_fel_convicted = round(mean(sent_prison_flag)*100,1),
            non_prison_of_fel_convicted = 100-prison_of_fel_convicted) %>%
  pivot_longer(cols = c(prison_of_fel_convicted, non_prison_of_fel_convicted), names_to = "variable") %>%
      rename(Total = value))%>%
  select(var, Total, `Asian/PI`, "Black", "Hispanic", "White")

) %>%
  
bind_rows(
#prison sentence length out of cases sentenced to prison
df %>%
  filter(sent_prison_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(race) %>%
  summarise(prison_length_yrs = round(mean(prison_length_days, na.rm = T)/365.25, 1)) %>%
  pivot_longer(cols = prison_length_yrs, names_to = "var") %>%
  pivot_wider(names_from = race, values_from = value)%>%
  bind_cols(total = df %>%
    filter(sent_prison_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
      summarise(prison_length_yrs = round(mean(prison_length_days, na.rm = T)/365.25, 1)) %>%
  pivot_longer(cols = prison_length_yrs, names_to = "variable") %>%
      rename(Total = value))%>%
  select(var, Total, `Asian/PI`, "Black", "Hispanic", "White")
) %>%
  

bind_rows(
  
  #arrest offense type
  df %>%
  filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(race, arrest_f_offense_type) %>%
  summarise(n = n()) %>%
  group_by(race) %>%
  mutate(percent = round(n/sum(n)*100,1)) %>%
  ungroup() %>%
  select(-n) %>%
  pivot_wider(names_from = race, values_from = percent) %>%
  rename(var = arrest_f_offense_type) %>%
  bind_cols(df %>%
    filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
      group_by(arrest_f_offense_type) %>%
      summarise(n = n()) %>%
      ungroup() %>%
      mutate(percent = round(n/sum(n)*100, 1)) %>%
      select(-n) %>%
      rename(Total = percent)) %>%
    select(-arrest_f_offense_type)
  
) %>%
  
bind_rows(
#average arrest hierarchy
df %>%
  filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(race) %>%
  summarise(avg_arrest_hierarchy = round(mean(max_arrest_hier_scaled), 4)) %>%
  pivot_longer(cols = avg_arrest_hierarchy, names_to = "var") %>%
  pivot_wider(names_from = race, values_from = value)%>%
  bind_cols(total = df %>%
    filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
      summarise(avg_arrest_hierarchy = round(mean(max_arrest_hier_scaled), 4)) %>%
  pivot_longer(cols = avg_arrest_hierarchy, names_to = "variable") %>%
      rename(Total = value))%>%
  select(var, Total, `Asian/PI`, "Black", "Hispanic", "White")
) %>%
  
bind_rows(
  #prior record category
   df %>%
  filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(race, prior_record_cat) %>%
  summarise(n = n()) %>%
  group_by(race) %>%
  mutate(percent = round(n/sum(n)*100,1)) %>%
  ungroup() %>%
  select(-n) %>%
  pivot_wider(names_from = race, values_from = percent) %>%
  rename(var = prior_record_cat) %>%
  bind_cols(df %>%
    filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
      group_by(prior_record_cat) %>%
      summarise(n = n()) %>%
      ungroup() %>%
      mutate(percent = round(n/sum(n)*100, 1)) %>%
      select(-n) %>%
      rename(Total = percent)) %>%
    select(-prior_record_cat)
  
) %>%
  
bind_rows(
  #gender
     df %>%
  filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(race, gender) %>%
  summarise(n = n()) %>%
  group_by(race) %>%
  mutate(percent = round(n/sum(n)*100,1)) %>%
  ungroup() %>%
  select(-n) %>%
  pivot_wider(names_from = race, values_from = percent) %>%
  rename(var = gender) %>%
  bind_cols(df %>%
    filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
      group_by(gender) %>%
      summarise(n = n()) %>%
      ungroup() %>%
      mutate(percent = round(n/sum(n)*100, 1)) %>%
      select(-n) %>%
      rename(Total = percent)) %>%
    select(-gender)
  
) %>%

bind_rows(
  #average age
  df %>%
  filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(race) %>%
  summarise(avg_age = round(mean(age), 1)) %>%
  pivot_longer(cols = avg_age, names_to = "var") %>%
  pivot_wider(names_from = race, values_from = value)%>%
  bind_cols(total = df %>%
    filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
      summarise(avg_age = round(mean(age), 1)) %>%
  pivot_longer(cols = avg_age, names_to = "variable") %>%
      rename(Total = value))%>%
  select(var, Total, `Asian/PI`, "Black", "Hispanic", "White")
  
  
) %>%
  
bind_rows(
  #number of cases
  df %>%
  filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(race) %>%
  summarise(n = n()) %>%
  pivot_longer(cols = n, names_to = "var") %>%
  pivot_wider(names_from = race, values_from = value)%>%
  bind_cols(total = df %>%
    filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
      summarise(n = n()) %>%
  pivot_longer(cols = n, names_to = "variable") %>%
      rename(Total = value))%>%
  select(var, Total, `Asian/PI`, "Black", "Hispanic", "White")
) %>%
  
  kbl() %>%
  kable_classic() %>%
  save_kable(file = "appendix table B1.html")
  

df %>%
  filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F")) %>%
  group_by(race) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  mutate(percent = round(n/sum(n)*100, 1))
  

```


#waffle charts
```{r waffles}

  #savings <- c(
 #   `Mortgage\n($84,911)` = 84911, `Auto and\ntuition loans\n($14,414)` = 14414,
 #   `Home equity loans\n($10,062)` = 10062, `Credit Cards\n($8,565)` = 8565
  #)
 # waffle(
   # savings / 392, rows = 7, size = 0.5, legend_pos = "bottom",
   # colors = c("#c7d4b6", "#a3aabd", "#a0d0de", "#97b5cf")
  #)
df %>%
  filter(court_disp ==1) %>%
  count(prior_record_cat) %>%
  mutate(percent = round(n/sum(n)*100, 1))

priors <- c(
  'No prior convictions\n(20.7%)' = 21, 'Prior conviction with no prior incarceration\n(17.0%)'= 17, 'Prior jail with no prior prison\n(37.2%)' = 37, 'Prior prison\n(25.1%)' = 25
)
waffle(
  priors , rows = 5, legend_pos = "top", colors = c("#2b83ba", "#5e3c99", "#d7191c", "#a6d96a")
)

df %>%
  filter(court_disp ==1) %>%
  count(arrest_f_offense_type) %>%
  mutate(percent = round(n/sum(n)*100, 1))


offtype <- c(
  'Violent\n(33.9%)' = 34, 'Property\n(32.2%)'= 32, 'Drug\n(12.1%)' = 12, 'Other\n(21.9%)' = 22
)
waffle(
  offtype , rows = 5, legend_pos = "top", 
  colors = c("#d7191c", "#2b83ba", "#a6d96a", "#5e3c99")
)




```
# Regressions

Set desired base levels for regressions
```{r message=FALSE, warning=FALSE}
df$race <- relevel(df$race, ref="White")
df$gender <- relevel(df$gender, ref="M")
df$arrest_f_offense_type <- as.factor(df$arrest_f_offense_type)
df$arrest_f_offense_type <- relevel(df$arrest_f_offense_type, ref="Property")
df$county <- as.factor(df$county)
df$county <- relevel(df$county, ref ="LOS ANGELES")
```

Conviction regressions
```{r message=FALSE, warning=FALSE}


convic <- glm(convicted_flag ~  days_prior_prison + days_prior_jail + 
              prior_conviction_summ_f_violent_flag_count + prior_conviction_summ_m_violent_flag_count + 
              prior_conviction_summ_f_property_flag_count + prior_conviction_summ_m_property_flag_count + 
              prior_conviction_summ_f_drug_flag_count + prior_conviction_summ_m_drug_flag_count + 
              prior_conviction_sex_flag_count  + prior_conviction_dv_flag_count +
              on_prob + prior_max_conv_hier_scaled +
              court_summ_f_violent_flag + court_summ_m_violent_flag +
              court_summ_f_property_flag + court_summ_m_property_flag + 
              court_summ_f_drug_flag + court_summ_m_drug_flag +
              court_sex_flag + court_dv_flag + max_court_hier_scaled +
              filed_mcharge_count + filed_fcharge_count  + 
              age + gender + race + 
              county, 
            data = df %>%
                  filter(filed_flag == 1,
                         gender %in% c("M", "F"),
                         race %in% c("Black", "White", "Hispanic", "Asian/PI")), 
            family = 'poisson')

# summary(conv_county)
conv <- tidy(coeftest(convic, vcov. = sandwich)) %>%
  mutate(significance = case_when(p.value < 0.001 ~ "***",
                                  p.value < 0.01 ~ "**",
                                  p.value < 0.05 ~ "*",
                                  p.value < 0.1 ~ ".",
                                  T ~ ""),
         relative_risk = round(exp(estimate), 4))

conv

#number of observations in regression
nrow(df %>%
        filter(filed_flag == 1,
               gender %in% c("M", "F"),
               race %in% c("Black", "White", "Hispanic", "Asian/PI")))
```

Felony conviction regressions
```{r message=FALSE, warning=FALSE}
#conviction regression, kitchen sink
conv_fel <- glm(conviction_felony_flag ~  days_prior_prison + days_prior_jail + 
                prior_conviction_summ_f_violent_flag_count + prior_conviction_summ_m_violent_flag_count + 
                prior_conviction_summ_f_property_flag_count + prior_conviction_summ_m_property_flag_count + 
                prior_conviction_summ_f_drug_flag_count + prior_conviction_summ_m_drug_flag_count + 
                prior_conviction_sex_flag_count  + prior_conviction_dv_flag_count + 
                on_prob + prior_max_conv_hier_scaled + 
                court_summ_f_violent_flag + court_summ_m_violent_flag +
                court_summ_f_property_flag + court_summ_m_property_flag +
                court_summ_f_drug_flag + court_summ_m_drug_flag +
                court_sex_flag + court_dv_flag + max_court_hier_scaled +
                filed_mcharge_count + filed_fcharge_count  + 
                age + gender + race + 
                county, 
              data = df %>%
                  filter(convicted_flag == 1,
                         gender %in% c("M", "F"),
                         race %in% c("Black", "White", "Hispanic", "Asian/PI")), 
              family = 'poisson')

# summary(conv_county)
felcon <- tidy(coeftest(conv_fel, vcov. = sandwich)) %>%
  mutate(significance = case_when(p.value < 0.001 ~ "***",
                                  p.value < 0.01 ~ "**",
                                  p.value < 0.05 ~ "*",
                                  p.value < 0.1 ~ ".",
                                  T ~ ""),
         relative_risk = round(exp(estimate), 4))

felcon

#number of observations in regression
nrow( df %>%
         filter(convicted_flag == 1,
                gender %in% c("M", "F"),
                race %in% c("Black", "White", "Hispanic", "Asian/PI")))

```

Prison sentencing regressions
```{r message=FALSE, warning=FALSE}

#detailing priors by type
prison_sent <- glm(sent_prison_flag ~  days_prior_prison + days_prior_jail + 
                  prior_conviction_summ_f_violent_flag_count + prior_conviction_summ_m_violent_flag_count + 
                  prior_conviction_summ_f_property_flag_count + prior_conviction_summ_m_property_flag_count + 
                  prior_conviction_summ_f_drug_flag_count + prior_conviction_summ_m_drug_flag_count + 
                  prior_conviction_sex_flag_count + prior_conviction_dv_flag_count + 
                  on_prob + prior_max_conv_hier_scaled + 
                  days_since_last_conv:prior_dummy +
                  conviction_summ_f_violent_flag + conviction_summ_m_violent_flag +
                  conviction_summ_f_property_flag + conviction_summ_m_property_flag +
                  conviction_summ_f_drug_flag + conviction_summ_m_drug_flag + 
                  conviction_sex_flag + conviction_dv_flag + max_conv_hier_scaled +
                  convicted_fcharge_count + convicted_mcharge_count  + 
                  age + gender + race + 
                  county, 
                data = df %>%
                  filter(conviction_felony_flag == 1,
                         gender %in% c("M", "F"),
                         race %in% c("Black", "White", "Hispanic", "Asian/PI")), 
                family = 'poisson')


pris_sent <- tidy(coeftest(prison_sent, vcov. = sandwich)) %>%
  mutate(significance = case_when(p.value < 0.001 ~ "***",
                                  p.value < 0.01 ~ "**",
                                  p.value < 0.05 ~ "*",
                                  p.value < 0.1 ~ ".",
                                  T ~ ""),
         relative_risk = round(exp(estimate), 4))

pris_sent

 # likelihood ratio test to compare models with and without race
#model_comp <- tidy(lrtest(prison_norace, prison_sent))


#plot_model(prison_sent, type = "pred", terms = c("race", "gender")) 

#number of observations in regression
nrow(df %>%
        filter(conviction_felony_flag == 1,
               gender %in% c("M", "F"),
               race %in% c("Black", "White", "Hispanic", "Asian/PI")))

```

Prison sentence length regressions
```{r message=FALSE, warning=FALSE}

prison_length <- lm(prison_length_days ~  years_prior_prison + years_prior_jail +
                      prior_conviction_summ_f_violent_flag_count + prior_conviction_summ_f_property_flag_count +  
                      prior_conviction_summ_f_drug_flag_count + prior_conviction_summ_m_violent_flag_count + 
                      prior_conviction_summ_m_property_flag_count +  prior_conviction_summ_m_drug_flag_count +  
                      prior_conviction_summ_f_other_sex_flag_count +  prior_conviction_summ_m_other_sex_flag_count + 
                      prior_conviction_summ_f_other_flag_count +  prior_conviction_summ_m_other_flag_count +
                      prior_conviction_sex_flag_count  + prior_conviction_dv_flag_count + prior_conviction_dui_flag_count +
                      on_prob + prior_max_conv_hier_scaled + 
                      inv_yrs_since_prior_conv +
                      conviction_summ_f_violent_flag + conviction_summ_m_violent_flag +
                      conviction_summ_f_property_flag + conviction_summ_m_property_flag +
                      conviction_summ_f_drug_flag + conviction_summ_m_drug_flag + 
                      conviction_summ_f_other_sex_flag + conviction_summ_m_other_sex_flag + 
                      conviction_summ_f_other_flag + conviction_summ_m_other_flag + 
                      conviction_sex_flag  + conviction_dv_flag + conviction_dui_flag +
                      conviction_violent_felony_flag + conviction_serious_felony_flag +
                      max_conv_hier_scaled +
                      convicted_fcharge_count + convicted_mcharge_count + 
                      age  + gender + race + 
                      county, 
                    data =  df %>%
                      filter(sent_prison_flag == 1,
                             gender %in% c("M", "F"),
                             race %in% c("Black", "White", "Hispanic", "Asian/PI")))

summary(prison_length)
tidy(coeftest(prison_length, vcov. = sandwich))
tidy(confint(prison_length, vcov. = sandwich))

df %>%
  count(conviction_summ_f_violent_flag, conviction_violent_felony_flag, conviction_serious_felony_flag)

df %>%
  select(matches("^conviction_.*_flag"))

df %>%
  count(conviction_summ_m_other_flag)

#lrtest(prison_length, prison_length_norace)
#anova(prison_length, prison_length_norace)
#what is the difference between lrtest and anova? which is appropriate to use here?


# plot_model(prison_length, type = "pred", terms = c("race"))


#number of observations in regression
nrow(df %>%
  filter(sent_prison_flag == 1, 
         gender %in% c("M", "F"),
         race %in% c("Black", "White", "Hispanic", "Asian/PI")))


summary(prison_length)$r.squared

prison_length_norace <- lm(prison_length_days ~  days_prior_prison + days_prior_jail +
                      prior_conviction_summ_f_violent_flag_count + prior_conviction_summ_f_property_flag_count +  
                      prior_conviction_summ_f_drug_flag_count + prior_conviction_summ_m_violent_flag_count + 
                      prior_conviction_summ_m_property_flag_count +  prior_conviction_summ_m_drug_flag_count +  
                      prior_conviction_sex_flag_count  + prior_conviction_dv_flag_count + 
                      on_prob + prior_max_conv_hier_scaled + 
                      days_since_last_conv:prior_dummy +
                      conviction_summ_f_violent_flag + conviction_summ_m_violent_flag +
                      conviction_summ_f_property_flag + conviction_summ_m_property_flag +
                      conviction_summ_f_drug_flag + conviction_summ_m_drug_flag + 
                      conviction_sex_flag  + conviction_dv_flag + max_conv_hier_scaled +
                      convicted_fcharge_count + convicted_mcharge_count + 
                      age  + gender + 
                      county, 
                    data =  df %>%
                      filter(sent_prison_flag == 1,
                             gender %in% c("M", "F"),
                             race %in% c("Black", "White", "Hispanic", "Asian/PI")))

tidy(lrtest(prison_length_norace, prison_length))
```

Variable tracking
```{r echo=FALSE, message=FALSE, warning=FALSE}
# names(df %>%
#   select(contains("prior")))
# 
# #relevant variables to control for criminal history
# prior_vars <- c("prior_conviction_restrain_flag_count", "prior_conviction_dv_possible_flag_count",  
#                 "prior_conviction_summ_f_violent_flag_count", "prior_conviction_summ_f_property_flag_count", 
#                 "prior_conviction_summ_f_drug_flag_count","prior_conviction_summ_f_other_sex_flag_count",
#                 "prior_conviction_summ_f_other_flag_count","prior_conviction_summ_m_violent_flag_count",
#                 "prior_conviction_summ_m_property_flag_count", "prior_conviction_summ_m_drug_flag_count",
#                 "prior_conviction_summ_m_other_sex_flag_count", "prior_conviction_summ_m_other_flag_count",
#                 "days_prior_prison", "days_prior_jail", "prior_max_conv_hier_scaled") 
# 
# 
# names(df %>%
#   select(contains("conviction")))

```



# Regression output including variable contributions (by category)

Create function to get McFadden Pseudo R2 values for groups of variables in a glm regression
```{r echo=FALSE, message=FALSE, warning=FALSE}

mcfaddenpr2 <- function(d, dv, crim_hist, current_crime, age, gender, location, race){
  
  dataf <- d
  
  categories <- c("crim_hist", "current_crime", "age", "gender", "location", "race")
  
  cats <- c(rep("crim_hist", length(crim_hist)), 
            rep("current_crime", length(current_crime)), 
            rep("age", length(age)),
            rep("gender", length(gender)),
            rep("location", length(location)),
            "race")
  
  ivs <- tibble(variable = c(crim_hist, current_crime, age, gender, location, race), cat = cats)
  

  test_df <- lapply(c(1:length(cats)), function(i){
    test <- ivs %>%
      filter(cat != categories[i]) %>%
      mutate(missing = categories[i])
    return(test)
  }) %>% bind_rows %>%
    bind_rows(ivs %>% mutate(missing = "full"))

  eqn_test <- lapply(c(1:length(unique(test_df$missing))), function(i){
    vars <- unique(test_df$missing)
    missing_var <- vars[i]
    test <- test_df %>%
      filter(missing == missing_var) %>%
      select(variable)

    str_lst <- ""
    for(i in c(1:nrow(test))){

      str_lst <- if_else(str_lst == "", paste0(test$variable[i]), paste0(str_lst, " + ", test$variable[i]))
      }

    fun_str <- paste0("glm(", dv, " ~ ", str_lst, ", data = dataf, family = 'poisson')")


    test_mod <- eval(parse(text = fun_str))
    
    test_pr2 <- as.tibble(pR2(test_mod)[4])%>%
              mutate(model = missing_var)
    
    return(test_pr2)
  }) %>%
    bind_rows() %>%
    mutate(difference = value[model == "full"] - value)
  
  
  # recreate full model to return output
  
  str_lst = ""
  for(i in c(1:nrow(ivs))){

      str_lst <- if_else(str_lst == "", paste0(ivs$variable[i]), paste0(str_lst, " + ", ivs$variable[i]))
      }

  fun_str <- paste0("glm(", dv, " ~ ", str_lst, ", data = dataf, family = 'poisson')")
  full_mod <- eval(parse(text = fun_str))
  
  # tidy output for full model
  mod_tidy <- tidy(coeftest(full_mod, vcov. = sandwich)) %>%
      mutate(significance = case_when(p.value < 0.001 ~ "***",
                                  p.value < 0.01 ~ "**",
                                  p.value < 0.05 ~ "*",
                                  p.value < 0.1 ~ ".",
                                  T ~ ""),
         relative_risk = round(exp(estimate), 4))
  
  # create a model without race to test against full model
  vars_norace <- ivs %>%
    filter(cat!= "race") %>%
    select(variable)
  
  str_lst = ""
  for(i in c(1:nrow(vars_norace))){

      str_lst <- if_else(str_lst == "", paste0(vars_norace$variable[i]), paste0(str_lst, " + ", vars_norace$variable[i]))
      }

  fun_str <- paste0("glm(", dv, " ~ ", str_lst, ", data = dataf, family = 'poisson')")
  full_mod_norace <- eval(parse(text = fun_str))
  
  model_comp <- tidy(lrtest(full_mod_norace, full_mod))
  
  #return a list with the pr2 values, the full model tidy ouput, the number of observations used in the model, and a list of the variables with their categories
  
  return(list(pr2 = eqn_test, full_model = mod_tidy, n = nrow(dataf), vars = ivs, lr_race_v_norace = model_comp))
}
```

Create function to get R2 values for groups of variables in a lm regression
```{r echo=FALSE, message=FALSE, warning=FALSE}

contribution_r2 <- function(d, dv, crim_hist, current_crime, age, gender, location, race){
  
  dataf <- d
  
  categories <- c("crim_hist", "current_crime", "age", "gender", "location", "race")
  
  cats <- c(rep("crim_hist", length(crim_hist)), 
            rep("current_crime", length(current_crime)), 
            rep("age", length(age)),
            rep("gender", length(gender)),
            rep("location", length(location)),
            "race")
  
  ivs <- tibble(variable = c(crim_hist, current_crime, age, gender, location, race), cat = cats)
  
  test_df <- lapply(c(1:length(cats)), function(i){
    test <- ivs %>%
      filter(cat != categories[i]) %>%
      mutate(missing = categories[i])
    return(test)
  }) %>% bind_rows %>%
    bind_rows(ivs %>% mutate(missing = "full"))

  eqn_test <- lapply(c(1:length(unique(test_df$missing))), function(i){
    vars <- unique(test_df$missing)
    missing_var <- vars[i]
    test <- test_df %>%
      filter(missing == missing_var) %>%
      select(variable)

    str_lst <- ""
    for(i in c(1:nrow(test))){

      str_lst <- if_else(str_lst == "", paste0(test$variable[i]), paste0(str_lst, " + ", test$variable[i]))
      }

    fun_str <- paste0("lm(", dv, " ~ ", str_lst, ", data = dataf)")


    test_mod <- eval(parse(text = fun_str))
    test_r2 <- as.tibble(summary(test_mod)$r.squared)%>%
              mutate(model = missing_var)
    return(test_r2)
  }) %>%
    bind_rows() %>%
    mutate(difference = value[model == "full"] - value)
  
  # recreate full model to return ouput
  str_lst = ""
  for(i in c(1:nrow(ivs))){

      str_lst <- if_else(str_lst == "", paste0(ivs$variable[i]), paste0(str_lst, " + ", ivs$variable[i]))
      }

  fun_str <- paste0("lm(", dv, " ~ ", str_lst, ", data = dataf)")
  full_mod <- eval(parse(text = fun_str))
  
  # tidy the full model output
  mod_tidy <- tidy(coeftest(full_mod, vcov. = sandwich)) %>%
      mutate(significance = case_when(p.value < 0.001 ~ "***",
                                  p.value < 0.01 ~ "**",
                                  p.value < 0.05 ~ "*",
                                  p.value < 0.1 ~ ".",
                                  T ~ ""))
  
  # create a model without race to test against full model
  vars_norace <- ivs %>%
    filter(cat!= "race") %>%
    select(variable)
  
  str_lst = ""
  for(i in c(1:nrow(vars_norace))){

      str_lst <- if_else(str_lst == "", paste0(vars_norace$variable[i]), paste0(str_lst, " + ", vars_norace$variable[i]))
      }

  fun_str <- paste0("lm(", dv, " ~ ", str_lst, ", data = dataf)")
  full_mod_norace <- eval(parse(text = fun_str))
  
  model_comp <- tidy(lrtest(full_mod_norace, full_mod))
  
  #return r2 contributions, full model tidy output, the number of observations, a list of variables with categories, and a liklihood ratio test comparison of the full model vs the full model without race
  return(list(r2 = eqn_test, full_model = mod_tidy, n = nrow(dataf), vars = ivs, lr_race_vs_norace = model_comp))

}

```

Regression and pseudo R2 output for conviction
```{r}

conv_pr2 <- mcfaddenpr2(d = df %>%
              filter(filed_flag == 1,
                         gender %in% c("M", "F"),
                         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
                         age >= 18),
            dv = "convicted_flag",
            crim_hist = c("years_prior_prison", "years_prior_jail", "prior_sent_probation_flag_count",
            "prior_conviction_summ_f_violent_flag_count", "prior_conviction_summ_m_violent_flag_count", 
            "prior_conviction_summ_f_property_flag_count", "prior_conviction_summ_m_property_flag_count", 
            "prior_conviction_summ_f_drug_flag_count", "prior_conviction_summ_m_drug_flag_count", 
            "prior_conviction_summ_f_other_sex_flag_count", "prior_conviction_summ_m_other_sex_flag_count", 
            "prior_conviction_summ_f_other_flag_count", "prior_conviction_summ_m_other_flag_count",
            "prior_conviction_violent_felony_flag_count", "prior_conviction_serious_felony_flag_count",
            "prior_conviction_sex_flag_count", "prior_conviction_dv_flag_count", "prior_conviction_dui_flag_count",
            "on_prob", "prior_max_conv_hier_scaled", 
            "inv_yrs_since_prior_conv"),
            current_crime = c("court_summ_f_violent_flag", "court_summ_m_violent_flag", "court_summ_f_property_flag", 
            "court_summ_m_property_flag", "court_summ_f_drug_flag", "court_summ_m_drug_flag", 
            "court_summ_f_other_sex_flag", "court_summ_m_other_sex_flag", 
            "court_summ_f_other_flag", "court_summ_m_other_flag", 
            "court_violent_felony_flag", "court_serious_felony_flag",
            "court_sex_flag", "court_dv_flag", "court_dui_flag",
            "max_court_hier_scaled", "filed_fcharge_count", "filed_mcharge_count",
            "combined_cycles_count"),
            age = c("age"), 
            gender = c("gender"),
            location = c("county"),
            race = c("race"))

conv_pr2

conv_relrisk <- conv_pr2$full_model %>%
  select(term, relative_risk, p.value) %>%
  filter(grepl("race", term)) %>%
  rename(race = term) %>%
  mutate(race = gsub("race", "", race)) %>%
  bind_rows(tibble(race = "White", relative_risk = 1))

df %>%
  filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F"),
         age >= 18) %>%
  group_by(race) %>%
  summarise(conviction_rate = mean(convicted_flag)*100) %>%
  left_join(conv_relrisk) %>%
  mutate(estimated_white_rate = conviction_rate/relative_risk) %>%
  mutate(across(contains("rate"), ~round(.x, 1)))

#make it a chart
df %>%
  filter(filed_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F"),
         age >= 18) %>%
  group_by(race) %>%
  summarise(conviction_rate = mean(convicted_flag)*100) %>%
  left_join(conv_relrisk) %>%
  mutate(estimated_white_rate = conviction_rate/relative_risk) %>%
  mutate(across(contains("rate"), ~round(.x, 1))) %>%
  rename(observed = conviction_rate) %>%
  rename(estimated = estimated_white_rate) %>%
  pivot_longer(c(observed, estimated)) %>%
  rename(conviction_rate = value) %>%
  ggplot(aes(fill=name, y= conviction_rate, x=race)) + 
    geom_bar(position="dodge", stat="identity") +
    scale_color_brewer(palette = "Set1")+
    ylim(0, 100) +
    geom_text(aes(label=round(conviction_rate, 0)), position=position_dodge(width=0.9), vjust=-0.25)

ggsave("conviction rate relative risk chart.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")

#plot pr2 values
conv_pr2$pr2 %>%
  filter(model != "full") %>%
  rename(variables = model) %>%
  rename(contribution = difference) %>%
  ggplot(aes(fill=variables, y= contribution, x=variables)) + 
    geom_bar(position="dodge", stat="identity") +
    scale_color_brewer(palette = "Set1")

#ggsave("variable contribution - conviction.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")

conv_pr2$full_model %>%
  kbl() %>%
  kable_classic() %>%
  save_kable(file = "conviction regression output table")

conv_pr2$n

conv_pr2$vars
	
```



Regression and pseudo R2 output for felony conviction
```{r}

fel_pr2 <- mcfaddenpr2(d = df %>%
              filter(convicted_flag == 1,
                         gender %in% c("M", "F"),
                         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
                         age >= 18),
            dv = "conviction_felony_flag",
            crim_hist = c("years_prior_prison", "years_prior_jail", "prior_sent_probation_flag_count",
            "prior_conviction_summ_f_violent_flag_count", "prior_conviction_summ_m_violent_flag_count", 
            "prior_conviction_summ_f_property_flag_count", "prior_conviction_summ_m_property_flag_count", 
            "prior_conviction_summ_f_drug_flag_count", "prior_conviction_summ_m_drug_flag_count", 
            "prior_conviction_summ_f_other_sex_flag_count", "prior_conviction_summ_m_other_sex_flag_count", 
            "prior_conviction_summ_f_other_flag_count", "prior_conviction_summ_m_other_flag_count",
            "prior_conviction_violent_felony_flag_count", "prior_conviction_serious_felony_flag_count",
            "prior_conviction_sex_flag_count", "prior_conviction_dv_flag_count", "prior_conviction_dui_flag_count",
            "on_prob", "prior_max_conv_hier_scaled", 
            "inv_yrs_since_prior_conv"),
            current_crime = c("court_summ_f_violent_flag", "court_summ_m_violent_flag", "court_summ_f_property_flag", 
            "court_summ_m_property_flag", "court_summ_f_drug_flag", "court_summ_m_drug_flag",
            "court_summ_f_other_sex_flag", "court_summ_m_other_sex_flag", 
            "court_summ_f_other_flag", "court_summ_m_other_flag", 
            "court_violent_felony_flag", "court_serious_felony_flag",
            "court_sex_flag", "court_dv_flag", "court_dui_flag", 
            "max_court_hier_scaled", "filed_fcharge_count", "filed_mcharge_count",
            "combined_cycles_count"),
            age = c("age"), 
            gender = c("gender"),
            location = c("county"),
            race = c("race"))

fel_pr2

fel_relrisk <- fel_pr2$full_model %>%
  select(term, relative_risk, p.value) %>%
  filter(grepl("race", term)) %>%
  rename(race = term) %>%
  mutate(race = gsub("race", "", race)) %>%
  bind_rows(tibble(race = "White", relative_risk = 1))

df %>%
  filter(convicted_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F"),
         age >= 18) %>%
  group_by(race) %>%
  summarise(felony_conviction_rate = mean(conviction_felony_flag)*100) %>%
  left_join(fel_relrisk) %>%
  mutate(estimated_white_rate = felony_conviction_rate/relative_risk) %>%
  mutate(across(contains("rate"), ~round(.x, 1)))

#chart
fel_pr2$pr2 %>%
  filter(model != "full") %>%
  rename(variables = model) %>%
  rename(contribution = difference) %>%
  ggplot(aes(fill=variables, y= contribution, x=variables)) + 
    geom_bar(position="dodge", stat="identity") +
    scale_color_brewer(palette = "Set1")

#ggsave("variable contribution - felony conviction.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")

fel_pr2$full_model %>%
  kbl() %>%
  kable_classic() %>%
  save_kable(file = "felony conviction regression output table.html")

fel_pr2$n
```


Regression and pseudo R2 output for prison sentencing
```{r}

sent_pr2 <- mcfaddenpr2(d = df %>%
              filter(conviction_felony_flag == 1,
                         gender %in% c("M", "F"),
                         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
                         age >= 18),
            dv = "sent_prison_flag",
            crim_hist = c("years_prior_prison", "years_prior_jail", "prior_sent_probation_flag_count",
            "prior_conviction_summ_f_violent_flag_count", "prior_conviction_summ_m_violent_flag_count", 
            "prior_conviction_summ_f_property_flag_count", "prior_conviction_summ_m_property_flag_count", 
            "prior_conviction_summ_f_drug_flag_count", "prior_conviction_summ_m_drug_flag_count", 
            "prior_conviction_summ_f_other_sex_flag_count", "prior_conviction_summ_m_other_sex_flag_count", 
            "prior_conviction_summ_f_other_flag_count", "prior_conviction_summ_m_other_flag_count",
            "prior_conviction_violent_felony_flag_count", "prior_conviction_serious_felony_flag_count",
            "prior_conviction_sex_flag_count", "prior_conviction_dv_flag_count", "prior_conviction_dui_flag_count",
            "on_prob", "prior_max_conv_hier_scaled", 
            "inv_yrs_since_prior_conv"),
            current_crime = c("conviction_summ_f_violent_flag", "conviction_summ_m_violent_flag", "conviction_summ_f_property_flag", 
            "conviction_summ_m_property_flag", "conviction_summ_f_drug_flag", "conviction_summ_m_drug_flag",
            "conviction_summ_f_other_sex_flag", "conviction_summ_m_other_sex_flag", 
            "conviction_summ_f_other_flag", "conviction_summ_m_other_flag", 
            "conviction_violent_felony_flag", "conviction_serious_felony_flag",
            "conviction_sex_flag", "conviction_dv_flag", "conviction_dui_flag",
            "max_conv_hier_scaled", "convicted_fcharge_count", "convicted_mcharge_count",
            "combined_cycles_count"),
            age = c("age"), 
            gender = c("gender"),
            location = c("county"),
            race = c("race"))

sent_pr2

pris_relrisk <- sent_pr2$full_model %>%
  select(term, relative_risk, p.value) %>%
  filter(grepl("race", term)) %>%
  rename(race = term) %>%
  mutate(race = gsub("race", "", race)) %>%
  bind_rows(tibble(race = "White", relative_risk = 1))

df %>%
  filter(conviction_felony_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F"),
         age >= 18) %>%
  group_by(race) %>%
  summarise(prison_rate = mean(sent_prison_flag)*100) %>%
  left_join(pris_relrisk) %>%
  mutate(estimated_white_rate = prison_rate/relative_risk) %>%
  mutate(across(contains("rate"), ~round(.x, 1)))

#chart
sent_pr2$pr2 %>%
  filter(model != "full") %>%
  rename(variables = model) %>%
  rename(contribution = difference) %>%
  ggplot(aes(fill=variables, y= contribution, x=variables)) + 
    geom_bar(position="dodge", stat="identity") +
    scale_color_brewer(palette = "Set1")

#ggsave("variable contribution - prison sentence.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")

sent_pr2$full_model %>%
  kbl() %>%
  kable_classic() %>%
  save_kable(file = "prison sentencing regression output table.html")

sent_pr2$n
```

R2 for sentence length
```{r}

length_r2 <- contribution_r2(d = df %>%
              filter(sent_prison_flag == 1,
                         gender %in% c("M", "F"),
                         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
                         age >= 18) ,
            dv = "prison_length_days",
            crim_hist = c("years_prior_prison", "years_prior_jail", "prior_sent_probation_flag_count",
            "prior_conviction_summ_f_violent_flag_count", "prior_conviction_summ_m_violent_flag_count", 
            "prior_conviction_summ_f_property_flag_count", "prior_conviction_summ_m_property_flag_count", 
            "prior_conviction_summ_f_drug_flag_count", "prior_conviction_summ_m_drug_flag_count", 
            "prior_conviction_summ_f_other_sex_flag_count", "prior_conviction_summ_m_other_sex_flag_count", 
            "prior_conviction_summ_f_other_flag_count", "prior_conviction_summ_m_other_flag_count", 
            "prior_conviction_violent_felony_flag_count", "prior_conviction_serious_felony_flag_count",
            "prior_conviction_sex_flag_count", "prior_conviction_dv_flag_count", "prior_conviction_dui_flag_count",
            "on_prob", "prior_max_conv_hier_scaled", 
            "inv_yrs_since_prior_conv"),
            current_crime = c("conviction_summ_f_violent_flag", "conviction_summ_m_violent_flag", "conviction_summ_f_property_flag", 
            "conviction_summ_m_property_flag", "conviction_summ_f_drug_flag", "conviction_summ_m_drug_flag", 
            "conviction_summ_f_other_sex_flag", "conviction_summ_m_other_sex_flag", 
            "conviction_summ_f_other_flag", "conviction_summ_m_other_flag", 
            "conviction_violent_felony_flag", "conviction_serious_felony_flag",
            "conviction_sex_flag", "conviction_dv_flag", "conviction_dui_flag",  
            "max_conv_hier_scaled", "convicted_fcharge_count", "convicted_mcharge_count",
            "combined_cycles_count"),
            age = c("age"), 
            gender = c("gender"),
            location = c("county"),
            race = c("race"))


length_r2

pris_len <- length_r2$full_model %>%
  select(term, estimate, p.value) %>%
  filter(grepl("race", term)) %>%
  rename(race = term) %>%
  mutate(race = gsub("race", "", race)) %>%
  bind_rows(tibble(race = "White", estimate = 0))

df %>%
  filter(sent_prison_flag == 1,
         race %in% c("Black", "White", "Hispanic", "Asian/PI"),
         gender %in% c("M", "F"),
         age >= 18) %>%
  group_by(race) %>%
  summarise(prison_yrs = mean(prison_length_days/365.25, na.rm = T)) %>%
  left_join(pris_len) %>%
  mutate(estimated_white_yrs = prison_yrs-  estimate/365.25) %>%
  mutate(across(contains("yrs"), ~round(.x, 1)))


#don't show the estimated white chart bc race was not significant

#chart
length_r2$r2 %>%
  filter(model != "full") %>%
  rename(variables = model) %>%
  rename(contribution = difference) %>%
  ggplot(aes(fill=variables, y= contribution, x=variables)) + 
    geom_bar(position="dodge", stat="identity") +
    scale_color_brewer(palette = "Set1")

#ggsave("variable contribution - prison length.png", path = "G:/CrimJustice/Dispositions by Race and Ethnicity/Figures 2019")

length_r2$full_model %>%
  kbl() %>%
  kable_classic() %>%
  save_kable(file = "prison sentence length regression output table.html")

length_r2$n
```

TEST - Regression and pseudo R2 output for jail sentencing 
```{r}

jail_pr2 <- mcfaddenpr2(d = df %>%
              filter(conviction_felony_flag == 0,
                     conviction_misd_flag == 1,
                     #convicted_flag == 1,
                     sent_prison_flag == 0,
                     gender %in% c("M", "F"),
                     race %in% c("Black", "White", "Hispanic", "Asian/PI")),
            dv = "sent_jail_flag",
            crim_hist = c("years_prior_prison", "years_prior_jail", "prior_sent_probation_flag_count",
            "prior_conviction_summ_f_violent_flag_count", "prior_conviction_summ_m_violent_flag_count", 
            "prior_conviction_summ_f_property_flag_count", "prior_conviction_summ_m_property_flag_count", 
            "prior_conviction_summ_f_drug_flag_count", "prior_conviction_summ_m_drug_flag_count", 
            "prior_conviction_summ_f_other_sex_flag_count", "prior_conviction_summ_m_other_sex_flag_count", 
            "prior_conviction_summ_f_other_flag_count", "prior_conviction_summ_m_other_flag_count",
            "prior_conviction_violent_felony_flag_count", "prior_conviction_serious_felony_flag_count",
            "prior_conviction_sex_flag_count", "prior_conviction_dv_flag_count", "prior_conviction_dui_flag_count",
            "on_prob", "prior_max_conv_hier_scaled", 
            "inv_yrs_since_prior_conv"),
            current_crime = c("conviction_summ_f_violent_flag", "conviction_summ_m_violent_flag", "conviction_summ_f_property_flag", 
            "conviction_summ_m_property_flag", "conviction_summ_f_drug_flag", "conviction_summ_m_drug_flag",
            "conviction_summ_f_other_sex_flag", "conviction_summ_m_other_sex_flag", 
            "conviction_summ_f_other_flag", "conviction_summ_m_other_flag", 
            "conviction_violent_felony_flag", "conviction_serious_felony_flag",
            "conviction_sex_flag", "conviction_dv_flag", "conviction_dui_flag",
            "max_conv_hier_scaled", "convicted_fcharge_count", "convicted_mcharge_count"),
            age = c("age"), 
            gender = c("gender"),
            location = c("county"),
            race = c("race"))

jail_pr2

df %>%
  mutate(orig_sent_incar = if_else(sent_prison_flag + sent_jail_flag + sent_imp_suspended_flag + sent_ex_suspended_flag > 0, 1, 0)) %>%
  count(orig_sent_incar, sent_prison_flag, sent_jail_flag, sent_imp_suspended_flag, sent_ex_suspended_flag, sent_suspended_flag)
```

TEST - R2 for jail length for misdemeanor convictions
```{r}

jaillength_r2 <- contribution_r2(d = df %>%
              filter(conviction_felony_flag == 0,
                     conviction_misd_flag == 1,
                     sent_prison_flag == 0,
                     sent_jail_flag == 1,
                     gender %in% c("M", "F"),
                     race %in% c("Black", "White", "Hispanic", "Asian/PI")) ,
            dv = "jail_length_days",
            crim_hist = c("years_prior_prison", "years_prior_jail", "prior_sent_probation_flag_count",
            "prior_conviction_summ_f_violent_flag_count", "prior_conviction_summ_m_violent_flag_count", 
            "prior_conviction_summ_f_property_flag_count", "prior_conviction_summ_m_property_flag_count", 
            "prior_conviction_summ_f_drug_flag_count", "prior_conviction_summ_m_drug_flag_count", 
            "prior_conviction_summ_f_other_sex_flag_count", "prior_conviction_summ_m_other_sex_flag_count", 
            "prior_conviction_summ_f_other_flag_count", "prior_conviction_summ_m_other_flag_count", 
            "prior_conviction_violent_felony_flag_count", "prior_conviction_serious_felony_flag_count",
            "prior_conviction_sex_flag_count", "prior_conviction_dv_flag_count", "prior_conviction_dui_flag_count",
            "on_prob", "prior_max_conv_hier_scaled", 
            "inv_yrs_since_prior_conv"),
            current_crime = c("conviction_summ_f_violent_flag", "conviction_summ_m_violent_flag", "conviction_summ_f_property_flag", 
            "conviction_summ_m_property_flag", "conviction_summ_f_drug_flag", "conviction_summ_m_drug_flag", 
            "conviction_summ_f_other_sex_flag", "conviction_summ_m_other_sex_flag", 
            "conviction_summ_f_other_flag", "conviction_summ_m_other_flag", 
            "conviction_violent_felony_flag", "conviction_serious_felony_flag",
            "conviction_sex_flag", "conviction_dv_flag", "conviction_dui_flag",  
            "max_conv_hier_scaled", "convicted_fcharge_count", "convicted_mcharge_count"),
            age = c("age"), 
            gender = c("gender"),
            location = c("county"),
            race = c("race"))


jaillength_r2


```

TEST - Regression and pseudo R2 output for suspension of incarcerative sentence
```{r}

susp_pr2 <- mcfaddenpr2(d = df %>%
              mutate(orig_sent_incar = if_else(sent_prison_flag + sent_jail_flag + sent_ex_suspended_flag > 0, 1, 0)) %>%
              filter(orig_sent_incar == 1,
                     gender %in% c("M", "F"),
                     race %in% c("Black", "White", "Hispanic", "Asian/PI")),
            dv = "sent_ex_suspended_flag",
            crim_hist = c("years_prior_prison", "years_prior_jail", "prior_sent_probation_flag_count",
            "prior_conviction_summ_f_violent_flag_count", "prior_conviction_summ_m_violent_flag_count", 
            "prior_conviction_summ_f_property_flag_count", "prior_conviction_summ_m_property_flag_count", 
            "prior_conviction_summ_f_drug_flag_count", "prior_conviction_summ_m_drug_flag_count", 
            "prior_conviction_summ_f_other_sex_flag_count", "prior_conviction_summ_m_other_sex_flag_count", 
            "prior_conviction_summ_f_other_flag_count", "prior_conviction_summ_m_other_flag_count",
            "prior_conviction_violent_felony_flag_count", "prior_conviction_serious_felony_flag_count",
            "prior_conviction_sex_flag_count", "prior_conviction_dv_flag_count", "prior_conviction_dui_flag_count",
            "on_prob", "prior_max_conv_hier_scaled", 
            "inv_yrs_since_prior_conv"),
            current_crime = c("conviction_summ_f_violent_flag", "conviction_summ_m_violent_flag", "conviction_summ_f_property_flag", 
            "conviction_summ_m_property_flag", "conviction_summ_f_drug_flag", "conviction_summ_m_drug_flag",
            "conviction_summ_f_other_sex_flag", "conviction_summ_m_other_sex_flag", 
            "conviction_summ_f_other_flag", "conviction_summ_m_other_flag", 
            "conviction_violent_felony_flag", "conviction_serious_felony_flag",
            "conviction_sex_flag", "conviction_dv_flag", "conviction_dui_flag",
            "max_conv_hier_scaled", "convicted_fcharge_count", "convicted_mcharge_count"),
            age = c("age"), 
            gender = c("gender"),
            location = c("county"),
            race = c("race"))

susp_pr2

```


